---
title: "Getting Started with JumpDiffSim"
author: "Kennedy Kayaki, Oh Dohyun, Ju Seong Nyeon, Lee Se Eun, Choi Jiwoo, Yuri Shin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with JumpDiffSim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  echo     = TRUE,
  message  = FALSE,
  warning  = FALSE,
  fig.width  = 6,
  fig.height = 4
)
```

## 1. Why Jump-Diffusion?

Standard Geometric Brownian Motion (GBM) assumes that log-returns are
normally distributed. Empirical equity and cryptocurrency returns, however,
exhibit well-documented departures from normality:

- **Leptokurtosis** (excess kurtosis > 0): more mass in the centre and tails
  than a Normal distribution predicts.
- **Negative skewness**: large negative moves are more common and more severe
  than large positive moves.

The **Merton (1976)** jump-diffusion model adds a compound Poisson process
to GBM, allowing the price to jump discontinuously at random times:

$$
dS_t = S_t \left[ (\mu - \lambda \bar{\mu}_J)\,dt +
       \sigma\,dW_t + dJ_t \right]
$$

where $J_t = \sum_{i=1}^{N(t)}(Y_i - 1)$, $N(t) \sim \text{Poisson}(\lambda t)$,
and $\log Y_i \sim \mathcal{N}(\mu_J, \sigma_J^2)$.

**JumpDiffSim** provides a clean, unified S4 interface for simulating paths
and calibrating parameters under this model, with all examples running
entirely offline.

---

## 2. Simulate Paths

```{r simulate, fig.width=6, fig.height=4}
library(JumpDiffSim)

# Create a MertonModel S4 object with default parameters
m <- MertonModel(
  mu      =  0.05,   # drift
  sigma   =  0.20,   # diffusion volatility
  lambda  =  1,      # average jumps per year
  mu_j    = -0.10,   # mean log-jump size
  sigma_j =  0.15    # std dev of log-jumps
)

# Display the model
show(m)

# Simulate 200 paths over 1 year with 252 daily steps
sim <- simulateMerton(m, n = 200, T_ = 1, steps = 252, seed = 42)

# Diagnostic plots
plts <- diagnosticPlots(sim)
print(plts$fan_chart)
print(plts$density)
```

The fan chart shows the 5th, 25th, 50th, 75th, and 95th percentile bands
of the 200 simulated paths. The density plot overlays the empirical
return histogram against a Normal curve, illustrating the heavy-tailed
character of Merton returns.

---

## 3. Fit Model to Data

```{r fit}
# Generate reproducible synthetic log-returns
# (all examples use jdSampleData() -- no internet required)
ret <- jdSampleData("merton", n = 500, seed = 42)

# Fit the Merton model via Maximum Likelihood Estimation
fit <- fitMerton(ret, verbose = FALSE)

# Parameter estimates and convergence
print(fit)

# 95% Wald confidence intervals
confint(fit)
```

The `fitMerton()` function uses L-BFGS-B optimisation and computes
Hessian-based standard errors via `numDeriv::hessian()`. The `converged`
slot confirms whether the optimiser reached a solution.

---

## 4. Interpreting Parameters

| Parameter | Meaning                             |
|-----------|-------------------------------------|
| mu        | Drift (expected continuous return)  |
| sigma     | Diffusion volatility                |
| lambda    | Average jumps per year              |
| mu_j      | Average log-size of each jump       |
| sigma_j   | Std deviation of jump sizes         |

A **negative** `mu_j` combined with a positive `lambda` implies that jumps
on average reduce the asset price — consistent with the crash-risk
interpretation of Merton (1976).

---

## 5. Theoretical Moments

```{r moments}
# Theoretical mean, variance, skewness, and excess kurtosis
# contributed by the jump component
jumpMoments(m)
```

Excess kurtosis greater than zero confirms that the Merton model generates
heavier tails than GBM, which is the key motivation for using a
jump-diffusion specification.

---

## References

- Merton, R.C. (1976). Option pricing when underlying stock returns are
  discontinuous. *Journal of Financial Economics*, 3(1-2), 125-144.
- Wickham, H. and Bryan, J. (2023). *R Packages* (2nd ed.). O'Reilly Media.
  <https://r-pkgs.org/>

